### 2022-04-19 ms/template.tex
reorder sectios

\usepackage{setspace}
\usepackage{lineno} % line numbering
\linenumbers

% Line spacing
\usepackage{setspace}
%\doublespacing

\usepackage{geometry}
\geometry{top=2.5cm, bottom=3cm, left=4.2cm, right=4.2cm}

% Fonts of the main text
\usepackage{fourier}
% Helvetica for sans serif fonts
\renewcommand{\sfdefault}{phv}

\usepackage[font={small,sf, doublespacing}, labelfont=bf]{caption}
\captionsetup{format=hang, font={it}, labelfont={bf}} % Format of captions
\captionsetup[table]{position=bottom}   %% or below