### 2022-07-04 ms/figs.Rmd
format supp table

---
title: "The French Covid-19 vaccination policy did not solve vaccination inequities: a nationwide longitudinal study on 64.5 million individuals"
author: "F. Débarre, E. Lecoeur, L. Guimier, M. Jauffret-Roustide, A.-S. Jannot"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    includes:
      in_header: template.tex
    fig_caption: yes
    number_sections: no
    toc: no
  bookdown::html_document2:
    self_contained: no
    keep_md: true
    number_sections: no
  bookdown::word_document2: 
    number_sections: no
    toc: no
bibliography: VaccinationInequalities.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("mapsf")

load("../code/outsavevalues.RData")
load("../code/data_indicators.RData") 
```

<!-- Other data -->

```{r, echo = FALSE}
insee.indicators.names <- names(typesDataAll)[!is.element(names(typesDataAll), c("Geography", "Elections"))]

tb <- typesDataNoCorr[!is.element(names(typesDataNoCorr), c("Geography", "Elections"))]
output.INSEE <- ""
for(i in 1:length(tb)) output.INSEE <- paste0(output.INSEE, names(tb)[i], ": n = ", tb[i], "; ")
output.INSEE <- substr(output.INSEE, 1, nchar(output.INSEE)-2)
```

```{r loadData_outLR, echo = FALSE}
load("../data/predictors.RData")
load("../code/outLR.RData")

# Extract OR values for the most associated variables
tmp <- outLR[outLR$varPred == "Asselineau", ]
maxAsselineau <- tmp[order(tmp$OR.abs, decreasing = TRUE), ][1, ]

tmp <- outLR[outLR$varPred == "Unemployment_Benef", ]
maxUnemployment_Benef <- tmp[order(tmp$OR.abs, decreasing = TRUE), ][1, ]

tmp <- outLR[outLR$varPred == "Immigrant", ]
maxImmigrant <- tmp[order(tmp$OR.abs, decreasing = TRUE), ][1, ]

tmp <- outLR[outLR$varPred == "NO.SE", ]
maxNOSE <- tmp[order(tmp$OR.abs, decreasing = TRUE), ][1, ]

tmp <- outLR[outLR$varPred == "Overcrowding_rate", ]
maxOvercrowding_rate <- tmp[order(tmp$OR.abs, decreasing = TRUE), ][1, ]

roundOR <- 3 # Number of decimals for OR

```

```{r exportSuppTable, eval = FALSE}
# Done only once

# Extract only key columns
# and remove the lines corresponding to permutations
tmp <- outLR[outLR$permutation == FALSE, c("varPred", "typePredFull", "thedate", "OR", "OR.CI.min", "OR.CI.max")]
# Rename the columns
names(tmp) <- c("predictor", "class of predictor", "date", "OR", "OR CI min", "OR CI max")

# Save as csv
write.csv(tmp, file = "../ms/SuppTable_OR.csv", row.names = FALSE)

# Compute permutation values
qq <- 0.99 # quantile for the permutations
dates <- unique(outLR$thedate) # values of the dates
qperm <- rep(NA, length(dates)) # initialize output

for(idate in seq_along(dates)){
  thedate <- dates[idate]
  # Subset of the data for the permutation tests
  subT <- outLR[outLR$thedate == thedate & outLR$permutation == TRUE, ]
  # Compute quantile
  xq <- quantile(subT$OR.abs, probs = qq)
  qperm[idate] <- xq
}

tmp.perm <- data.frame(date = dates, q99 = qperm)
tmp.perm

write.csv(tmp.perm, file = "../ms/SuppTable_ORpermutations99.csv", row.names = FALSE)

# Combine the two in a single table
#  New column to store the result
tmp$significant <- NA
tmp$OR_permutation <- NA

for(i in seq_len(nrow(tmp))){
  # Sort CI values
  tmp[i, c("OR CI min", "OR CI max")] <- sort(tmp[i, c("OR CI min", "OR CI max")])
  # Test significativity
  ORabs <- max(tmp[i, "OR"], 1/tmp[i, "OR"])
  ORref <- tmp.perm[tmp.perm$date == tmp[i, "date"], "q99"]
  if(ORabs > ORref) tmp[i, "significant"] <- TRUE
  if(ORabs < ORref) tmp[i, "significant"] <- FALSE
  tmp[i, "OR_permutation"] <- ORref
}

# Round the values
dgt <- 3
tmp$OR <- round(tmp$OR, digits = dgt)
tmp$`OR CI min` <- round(tmp$`OR CI min`, digits = dgt)
tmp$`OR CI max` <- round(tmp$`OR CI max`, digits = dgt)

# Save as csv
write.csv(tmp, file = "../ms/SuppTable_OR.csv", row.names = FALSE)

```

```{r figManhattan, echo=FALSE, fig.height=9, fig.width=10, fig.cap="Manhattan plots of the Odds ratios for each of the indicator of our dataset, by date. Left column: around the Sanitary Pass implementation; right column: around the Vaccine Pass implementation. The top odds ratios are labelled at each time point; the symbol next to the name indicates the direction of the effect. The gray rectangle corresponds to the 99% percentile of odds ratios in the permuted data; points falling in the rectangle are considered as non-significant."}
source("../code/2_plot-manhattan.R")
layout(matrix(1:6, ncol = 2))
# PLOT
plotManhattan(outLR, ntop = 5, renameLabels = TRUE)
```

```{r manhattanSingles, fig.height=5, fig.width=8}
plotManhattan(outLR, ntop = 5)
```

```{r figOverTime, fig.height=8, fig.width=4, fig.cap="Age-adjusted vaccination rates among adults, over time, by decile of each indicator (presented by a color gradient). The vertical lines indicate the dates of announcements and implementations of the sanitary and vaccine passes.", echo = FALSE}
load("../code/outTime.RData")
source("../code/2_plot-overTime.R")
layout(matrix(1:3, ncol = 1))
plotPropTime(outA, plotDates = TRUE, plotGraduations = TRUE)
```

```{r overTimeSingles, fig.height=4, fig.width=6}
plotPropTime(outA, plotDates = TRUE, plotGraduations = TRUE)
```

```{r figMap, fig.height=5, fig.width=4, fig.cap="Adult vaccination rates by district of mainland France", echo = FALSE}
source("../code/2_plot-map.R")
source("../code/1_functions-adjustAges.R")
load("../data/mapFiles_withDepReg.RData")
load("../data/mapchefslieux.RData")
# Choose date
thedate <- datesAnalysis[6]
# Vaccination data by age at that date
tmpVacc <- vacc[vacc$date == thedate & vacc$classe_age != "TOUT_AGE", ]
#tmpVacc2 <- vacc[vacc$date == thedate & vacc$classe_age == "TOUT_AGE", ]
#tmpVacc2$taux_cumu <- tmpVacc2$taux_cumu_1_inj

# Compute adult vaccination rate
adultVaccinationRate <- adultVacc(tmpVacc)

#plotMapVacc(adultVaccinationRate, thetitle = paste0("Adult vaccination rates on ", thedate), palName = "Hiroshige", byp = 5)

# Compute mean vaccination rate
meanVacc <- sum(adultVaccinationRate$n_cumu) / sum(adultVaccinationRate$population_carto)
# Compute difference to max to center the color scale
maxDiff <- max(c(max(adultVaccinationRate$taux_cumu) - meanVacc, meanVacc - min(adultVaccinationRate$taux_cumu)))

plotMapVacc(adultVaccinationRate, thetitle = paste0("Adult vaccination rates on ", thedate), palName = "Cassatt2", byp = 2, zmin = floor(100*(meanVacc - maxDiff)), zmax = ceiling(100*(meanVacc + maxDiff)))

#plotMapVacc(tmpVacc2, thetitle = paste0("Adult vaccination rates on ", date6))

```